Nail: A practical interface generator for data formats

### 2014-04-03 paper/relwk.tex
Fixed XXX

Parser generators for binary protocols were first introduced by the
Hammer~\cite{hammer-parser} parser.
While previous parser generators could also be used to write grammars
for binary protocols,\footnote{Theoretically speaking, the alphabet
over which a grammar is an  abstract set, so most algorithms work
just as well on an alphabet of $\{0,1\}$.} doing so is practically
inconvenient. Hammer allows the programmer to specify a grammar in
terms of bits and bytes instead of characters. Common concerns, such as
endianness and bit-packing are transparently hidden by the library. Hammer
implements grammars as language-integrated parser combinators, an approach
popularized by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
parser combinator style (to our knowledge, first  described
in~\cite{burge1975recursive}) is a natural way of concisely expressing
top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For
more background on the history of expressing grammars, see Bryan Ford's
masters thesis~\cite{ford2002packrat}, which also describes the default
parsing algorithm used by Hammer.} by composing them from one or multiple
sub-parsers.  Hammer then constructs a tree of function pointers which
can be invoked to parse a given input into an abstract syntax tree (AST).
\XXX[how does Nail relate to Hammer, in terms of what's new, what it
solves, etc?]

>>>>

Parser generators for binary protocols were first introduced by the
Hammer~\cite{hammer-parser} parser.
While previous parser generators could also be used to write grammars
for binary protocols,\footnote{Theoretically speaking, the alphabet
over which a grammar is an  abstract set, so most algorithms work
just as well on an alphabet of $\{0,1\}$.} doing so is practically
inconvenient. Hammer allows the programmer to specify a grammar in
terms of bits and bytes instead of characters. Common concerns, such as
endianness and bit-packing are transparently hidden by the library. Hammer
implements grammars as language-integrated parser combinators, an approach
popularized by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The
parser combinator style (to our knowledge, first  described
in~\cite{burge1975recursive}) is a natural way of concisely expressing
top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For
more background on the history of expressing grammars, see Bryan Ford's
masters thesis~\cite{ford2002packrat}, which also describes the default
parsing algorithm used by Hammer.} by composing them from one or multiple
sub-parsers.  Hammer then constructs a tree of function pointers which
can be invoked to parse a given input into an abstract syntax tree (AST).

Nail improves upon Hammer in three ways. First, Nail  generates output besides recognising input.
Second, Nail does not require the programmer to write potentially insecure semantic actions. Last,
Nail's structural dependencies allow it to work with protocols Hammer cannot recognise, such as
protocols with offset fields or length fields (Hammer has extremely limited support for length
fields. It can parse arrays immediately preceeded by their length).
